---
title: "IS-MRA: Simulation study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IS-MRA: Simulation study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
We start by simulating data. 
```{r setup}
library(ISMRAanalyses)
library(sp)
library(ISMRAanalyses)

# setwd("/home/luc/INLAMRAfiles/INLAMRApaper1/simulations")
setwd("/home/luc/ISMRAfiles/")
```
The simulated data are based on satellite imagery data obtained in Maharashtra. We first define a rectangular zone in the longitude-latitude projection. Since MODIS data are in the sinusoidal projection, we project the zone we just defined in that coordinate system.
```{r}
MaharashtraPolygonEdges <- rbind(c(18.8, 73.2), c(18.8, 73.4), c(18.6, 73.4), c(18.6, 73.2), c(18.8, 73.2))
MaharashtraPolygonEdges <- MaharashtraPolygonEdges[ , 2:1]
MaharashtraPolygon <- sp::SpatialPolygons(Srl = list(sp::Polygons(list(sp::Polygon(coords = MaharashtraPolygonEdges)), ID = "Mumbai")))
raster::crs(MaharashtraPolygon) <- sp::CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
MaharashtraPolygonOtherCRS <- sp::spTransform(MaharashtraPolygon, CRSobj = sp::CRS("+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +R=6371007.181 +units=m +no_defs"))
```
We then import land cover and elevation values from the MODIS and ASTER data files we obtained from EarthData.
```{r}
folderForEarthDataFiles <- "/store/luc/rawDataFiles/"
landCoverFiles <- list.files(folderForEarthDataFiles, pattern = "MCD*", full.names = TRUE)
landCoverRasterSinusoidal <- produceLandCover(landCoverFiles, regionPolygon = MaharashtraPolygonOtherCRS)
elevationFiles <- list.files(path = folderForEarthDataFiles, pattern = "*dem.tif", full.names = TRUE)
elevationRasterList <- lapply(elevationFiles, raster::raster)
```
From the imported data, we produce the covariate matrix used in the simulations.
```{r}
dayOffset <- 121 # Days are listed numerically, from 1 to 366 (leap year). Day 121 corresponds to April 30.
dayRange <- 27:29 # We extract data for May 27 until May 29.
collectionDatesPOSIX <- as.POSIXct(paste("2012-05-", dayRange, sep = ""))
covariateData <- prepareCovariateDataForISMRA(elevationRasterList, landCoverRasterSinusoidal, collectionDatesPOSIX)
saveRDS(covariateData, file = "data/covariatesForSimulationStudy.rds", compress = TRUE)
```
We are now ready to simulate $1,000$ vectors from a spatiotemporal Gaussian process.
```{r}
timepoints <- as.numeric(time(covariateData@time))/(3600*24) # Time is in days

# Parameter values based on real data analysis (cf. table 1 in paper)
# Spatial parameters
spSmoothness <- 1.5
spRange <- 5.66
spScale <- 4.14

# Time parameters
timeSmoothness <- 0.5
timeRange <- 3.60
timeScale <- 1

# Uncorrelated error
errorSD <- 0.5

covfun <- function(spacetimeObject, timepointsVec) {
  spatialDistances <- fields::rdist(spacetimeObject@sp@coords)
  timeDistances <- fields::rdist(timepointsVec)
  GPvecchia::MaternFun(spatialDistances, c(spScale^2, spRange, spSmoothness)) * GPvecchia::MaternFun(timeDistances, c(timeScale^2, timeRange, timeSmoothness))
}

covarStruct <- covfun(covariateData, timepoints) + errorSD^2 * diag(nrow(covariateData@data))

choleskyDecomp <- t(chol(covarStruct))

set.seed(10)

numReplicates <- 1000
fieldValuesWithError <- replicate(n = 1000, expr = as.numeric(choleskyDecomp %*% rnorm(nrow(covarStruct))))
```
We then add fixed effects to the field values, yielding the data we are going to model using the methods we selected for the simulation study.
```{r}
landCoverEffects <- c(
  Water = 0,
  EvergreenNeedleleaf = 1.09,
  EvergreenBroadleaf = 0.979,
  DeciduousBroadleaf = 1.181,
  MixedForest = 1.087,
  ClosedShrublands = 2.097,
  OpenShrublands = 1.448,
  WoodySavannas = 1.329,
  Savannas = 1.393,
  Grasslands = 1.506,
  PermanentWetlands = 0.705,
  Croplands = 1.588,
  Urban = 1.623,
  CroplandNaturalMosaics = 1.430,
  NonVegetated = 0.667)

# Water is the reference (landCover = 0)
referenceLandCover <- "Water"
referenceLandCoverIndex <- match(referenceLandCover, names(landCoverEffects))

timeEffects <- c(
  May27 = -0.033,
  May28 = -6.319,
  May29 = 1.155) + 0.033 # We're moving the reference day from May 25 to May 27, hence + 0.033
elevationEffect <- -0.001
interceptValue <- 40 # This is arbitrary. It should not affect the analyses.

landCoverFactor <- factor(covariateData@data$landCover, levels = 0:14, labels = names(landCoverEffects))
timepointsFactor <- factor(timepoints, levels = unique(timepoints), labels = names(timeEffects))

referenceTimepoint <- "May27"
referenceTimepointIndex <- match(referenceTimepoint, names(timeEffects))

covariateMatrix <- cbind(model.matrix(object = ~ landCover, data = data.frame(landCover = landCoverFactor)), elevation = covariateData@data$elevation, model.matrix(object = ~ time, data = data.frame(time = timepointsFactor))[ , -1]) # We remove the intercept, which has already been created for land cover. Ocean and May 27 are reference categories.
colnames(covariateMatrix) <- replace(colnames(covariateMatrix), which(colnames(covariateMatrix) == "(Intercept)"), "Intercept")
fixedEffectsVec <- c(Intercept = interceptValue, landCoverEffects[-referenceLandCoverIndex], elevation = elevationEffect, timeEffects[-referenceTimepointIndex])
colnames(covariateMatrix) <- gsub(pattern = "landCover", replacement = "", x = colnames(covariateMatrix))
colnames(covariateMatrix) <- gsub(pattern = "time", replacement = "", x = colnames(covariateMatrix))

# Getting the simulated values.

observedValues <- fieldValuesWithError + as.vector(covariateMatrix %*% fixedEffectsVec[colnames(covariateMatrix)])

# Fixing the covariate matrix
landCoverTypes <- names(landCoverEffects)[-referenceLandCoverIndex]
presentLandCoverTypes <- landCoverTypes[sapply(landCoverTypes, function(coverType) !all(covariateMatrix[ , coverType] == 0))]

covariateMatrixNoAbsentLandCovers <- cbind(covariateMatrix[ , -match(landCoverTypes, colnames(covariateMatrix))], covariateMatrix[ , presentLandCoverTypes])

simulatedDatasetsList <- list(
  responses = observedValues,
  covariates = covariateMatrixNoAbsentLandCovers[ , -match("Intercept", colnames(covariateMatrixNoAbsentLandCovers))],
  coords = covariateData@sp@coords,
  time = timepointsFactor,
  timeNumeric = timepoints)

saveRDS(simulatedDatasetsList, file = "data/dataForSimulations.rds", compress = TRUE)
```

As a sanity check, we can plot one of the simulated datasets.
```{r, eval=TRUE}
numRasterRows <- numRasterCols <- 48 # Approximate, hard to recover because data were collected in the sinusoidal projection, which does not translate to a perfectly-spaced grid of coordinates in the longitude/latitude projection.
observationsToPlot <- which(simulatedDatasetsList$time == "May28")
fields::quilt.plot(
  x = simulatedDatasetsList$coords[ , "x"][observationsToPlot],
  y = simulatedDatasetsList$coords[ , "y"][observationsToPlot],
  z = simulatedDatasetsList$responses[observationsToPlot],
  nx = numRasterRows, ny = numRasterCols)
```

